```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Upload needed libraries

library(DescTools)
library(ggplot2)
library(dplyr)
library(readxl)
library(readr)
library(purrr)
library(Hmisc)
library(pastecs)
library(ggpubr)
library(nortest)
library(tidyr)
library(reshape)
library(reporttools)
library(naniar)
library(corrplot)
library(corrgram)
library(qgraph)
library(networkD3)


rm(list=ls())
```
#Upload dataset

```{r updataset, include=FALSE}

#setwd("E:/Utenti/Andrea/Dropbox (modeFinance)/LAVORO/Andrea/W82-ProgettoUniPavia/Andrea")

final_dataset_smes <- read_csv("final_dataset_smes_and.csv")
data_complete <- final_dataset_smes[-1]


```
# Specify the type of variables

```{r type, echo=FALSE}
data_complete[,c(20:25)] <- apply(data_complete[,c(20:25)], 2, function(x) as.factor(as.character(x)))

data <- data_complete[,c(1:19,26, 27)]

```
# Summary of the dataset 
The function describe gives the sample size, missing, unique, mean, 5,10,25,50,75,90,95th percentiles, lowest, highest value for all columns in the dataset

```{r summary2, echo=FALSE}

colnames(data)[c(1:15)] = c("leverage", "fin_leverage", "ta_tl","current_ratio", "quick_ratio","assets_coverage","EBIT_interest", "ROI", "ROE","turnover_assets1","turnover_assets2", "interest_paid_weight", "EBITDA_int_paid", "EBITA_op_revenues","EBITA_sales")

describe(data)

```

# We notice DSO DPO problem:

```{r histDPO, echo=FALSE}
 ggplot(data, aes(DPO)) +
  geom_histogram(aes(y = stat(count/nrow(data))), colour="black")
 
ggplot(data, aes(DPO)) +
  geom_histogram( aes(y = stat(count/nrow(data))), colour="black",binwidth = 10 )

 ggplot(data, aes(DPO)) +
  geom_histogram(aes(y = stat(count/nrow(data))), colour="black",binwidth = 10)+ xlim(-10, 180) 
  
 ggplot(data, aes(DSO)) +
  geom_histogram( aes(y = stat(count/nrow(data))), colour="black")

ggplot(data, aes(DSO)) +
  geom_histogram(aes(y = stat(count/nrow(data))),colour="black",binwidth = 10)

 ggplot(data, aes(DSO)) +
  geom_histogram(aes(y = stat(count/nrow(data))),colour="black",binwidth = 10)+ xlim(-10, 180) 

```

# Short format FS 
issue of short form FS-> small Cos. Let's see what happens if we "remove" companies that file short form financials (i.e. turnover > 8.8 M)

```{r summary3, echo=FALSE}
data2 <- data[data$comp_size==2,]
data1 <- data[data$comp_size==1,]

dim(data2)
dim(data1)

#describe(data2)
#describe(data1)


 ggplot(data2, aes(DPO)) +
  geom_histogram(aes(y = stat(count/nrow(data))), colour="black",binwidth = 10)+ xlim(-10, 180) 
 ggplot(data1, aes(DPO)) +
  geom_histogram(aes(y = stat(count/nrow(data))), colour="black",binwidth = 10)+ xlim(-10, 180) 
  

 ggplot(data2, aes(DSO)) +
  geom_histogram(aes(y = stat(count/nrow(data))),colour="black",binwidth = 10)+ xlim(-10, 180) 
 ggplot(data1, aes(DSO)) +
  geom_histogram(aes(y = stat(count/nrow(data))),colour="black",binwidth = 10)+ xlim(-10, 180) 
```

#Size dependency of ratios
Let us see if there is different behaviour of other rations depending on size

```{r plots, echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

x <-  c(.05,.10, .25, .50, .75, .90)
dat <- data.frame(
    qunantiles = x,
    q_lev_2 = quantile(data2$leverage, x ),
    q_flev_2 =quantile(data2$fin_leverage, x),
    q_qr_2 =  quantile(data2$quick_ratio, x),
    q_cr_2 =  quantile(data2$current_ratio, x),
    q_roi_2 =  quantile(data2$ROI, x),
    q_roe_2 =  quantile(data2$ROE, x),
    q_lev = quantile(data1$leverage, x ),
    q_flev = quantile(data1$fin_leverage, x),
    q_qr = quantile(data1$quick_ratio, x),
    q_cr =  quantile(data1$current_ratio, x),
    q_roi =  quantile(data1$ROI, x),
    q_roe =  quantile(data1$ROE, x)
)


ggplot(dat, aes(x= qunantiles, y = leverage, color = variable)) + 
    geom_point(aes(y = q_lev_2, col = "large_corps")) + 
    geom_point(aes(y = q_lev, col = "small_corps"))  

ggplot(dat, aes(x= qunantiles, y = ROI, color = variable)) + 
    geom_point(aes(y = q_roi_2, col =  "large_corps")) + 
    geom_point(aes(y = q_roi, col = "small_corps"))

ggplot(dat, aes(x= qunantiles, y = ROE, color = variable)) + 
    geom_point(aes(y = q_roe_2, col = "large_corps")) + 
    geom_point(aes(y = q_roe, col = "small_corps"))

```
# Boxplot 
Another interesting tool to carry on data analysis are Box Plots. Once again 
knowledge of the facts is mandatory to carry on a reliable analaysis

```{r, out.width='1\\linewidth', fig.asp=0.5, fig.ncol = 1, fig.cap="output",fig.align = "center", echo=FALSE}
data %>%
  keep(is.numeric) %>% 
  boxplot(data, las=2)

```

#We've to keep in mind what we're looking at: ROI has a different range than leverage and both are far different than turnover... You cannot compare everything toghether.

```{r, box_plot_sector, echo=FALSE}

data$sector <- factor(data$sector) 
ggplot(data)+ geom_boxplot(aes(x = "leverage", y=data$leverage))+
 geom_boxplot(aes(x = "fin_leverage", y=data$fin_leverage))+ ylim(-5, 15) + theme(axis.title.y = element_blank()) 


ggplot(data)+ geom_boxplot(aes(x = "ROI", y=ROI))+
 geom_boxplot(aes(x = "ROE", y=data$ROE))+ ylim(-0.2, 1) + theme(axis.title.y = element_blank()) 

```

#Box plot come handy to study other dependencies, such as ratios vs sector, or size:

```{r, box_plot_sector_size, echo=FALSE}

data$sector <- factor(data$sector) 
bp <- ggplot(data, aes(x = sector, y=leverage, fill = sector))+ geom_boxplot()
bp + facet_grid(. ~  turnover > 8800)+ ylim(-10, 20)


bp <- ggplot(data, aes(x = sector, y=current_ratio, fill = sector))+ geom_boxplot()
bp + facet_grid(. ~  turnover > 8800)+ ylim(0, 3)


bp <- ggplot(data, aes(x = sector, y=quick_ratio, fill = sector))+ geom_boxplot()
bp + facet_grid(. ~  turnover > 8800) + ylim(0, 3) 


```

# Analysis Active vs Defaulted
In the end, the key point when developing a scoring model, is to identify ratio's discriminating capacity...

```{r plots_2, echo=FALSE}
data$status <- factor(data_complete$status)

ggplot(data, aes(x = status, y=leverage))+ geom_boxplot() + ylim(-10, 20)
ggplot(data, aes(x = status, y=leverage, fill = sector))+ geom_boxplot() + ylim(-10, 20)

ggplot(data, aes(x = status, y=ROI))+ geom_boxplot() + ylim(-0.2, 0.4)
ggplot(data, aes(x = status, y=ROI, fill = sector))+ geom_boxplot() + ylim(-0.2, 0.4)

ggplot(data, aes(x = status, y=ROE))+ geom_boxplot() + ylim(-0.2, 0.3)
ggplot(data, aes(x = status, y=ROE, fill = sector))+ geom_boxplot() + ylim(-0.2, 0.3)

```

# NOTICE
leverage of defaulted companies is lower than active ones? Weird... let's see what appens:

```{r plots_3, echo=FALSE}
x <-  c(.1,.2,.3,.4,.5,.6,.7,.8,.9)
dattt <- data.frame(
    quantiles_x = x,
    q_lev_a = quantile(data$leverage[data$status == 0], x ),
    q_lev_f = quantile(data$leverage[data$status == 1], x ),
    q_roi_a = quantile(data$ROI[data$status == 0], x ),
    q_roi_f = quantile(data$ROI[data$status == 1], x ),
    q_roe_a = quantile(data$ROE[data$status == 0], x ),
    q_roe_f = quantile(data$ROE[data$status == 1], x )
)


p1 <- ggplot(dattt, aes(x= quantiles_x, y = leverage, color = variable)) + 
    geom_point(aes(y = q_lev_a, col = "active")) + 
    geom_point(aes(y = q_lev_f, col = "defaulted")) 

p2 <- ggplot(dattt, aes(x= quantiles_x, y = ROI, color = variable)) + 
    geom_point(aes(y = q_roi_a, col = "active")) + 
    geom_point(aes(y = q_roi_f, col = "defaulted")) 

p3 <- ggplot(dattt, aes(x= quantiles_x, y = ROE, color = variable)) + 
    geom_point(aes(y = q_roe_a, col = "active")) + 
    geom_point(aes(y = q_roe_f, col = "defaulted")) 


multiplot(p1,p2,p3)
  
```

# Correlation analysis
Once identified companies with the highest discriminating capacity, we've to check for correlation: we may have three very "powerful" variables, but if they're totally corrleated, they do not bring value to our model.

```{r plots_4, echo=FALSE}

#dim(data)

data_test <- data[,c(1:18)] %>%
  keep(is.numeric) 
correlations <- cor(data_test)
corrplot(correlations, method="circle")

```

#sector subgroups

```{r plots_5, echo=FALSE}

#§data_test <- data %>%
 # keep(is.numeric) 
#correlations <- cor(data_test)
#corrplot(correlations, method="circle")

data_test <- data2[data2$sector == 4, ] %>%
  keep(is.numeric)
correlations <- cor(data_test[,c(4:5)])
corrplot(correlations, method="number",title= "Construtcions", mar=c(0,0,1,0))

data_test <- data2[data2$sector == 6, ] %>%
  keep(is.numeric)
correlations <- cor(data_test[,c(4:5)])
corrplot(correlations, method="number",title= "Services" , mar=c(0,0,1,0))


```

